Dynamic Species Interactions in the San Nicolas Island Kelp Forest

### 2018-06-13 EDM_SNI_analysis.Rmd
tailing exploratory analyses

# for black plots
plot_theme_black <-   theme_minimal()+
  theme(text=element_text(family="Rockwell",size=12,color="white"),
        line = element_line(color="gray40"),
        legend.text = element_text(size=14),
        axis.title=element_text(family="Rockwell",size=14),
        axis.text=element_text(family="Rockwell",color="white",size=8),
        strip.background = element_rect(colour="black"),
        panel.border = element_rect(color="white",fill=NA),
        panel.background = element_rect(fill="black"),
        plot.background = element_rect(fill="black"),
        panel.grid = element_line(color="gray40"),
        panel.grid.major = element_line(color="gray40"),
        panel.grid.minor = element_line(color="gray40"))
theme_set(plot_theme)
```

@@

```{r mean density over time, echo=F,message=F}
# Species naming key for use in plotting
namekey <- tibble(dataset=c(rep("Benthic density",6),rep("Physical",5)),short=names(westend[3:13]),
                  long=c("Laminaria","Macrocystis >1m","Pterygophora","Purple urchin","Red urchin","Macrocystis <1m","Multivariate ENSO Index","Pacific Decadal Oscillation","North Pacific Gyre Oscillation","Significant Wave Height","Sea Surface Temperature"),plotting=c("L. far","M. pyr","P. cal","S. pur","M. fra","M. pyr (j)","(P) MEI","(P) PDO","(P) NPGO","(P) Max Hs","(P) SST"))

@@

# position adjustment for visual clarity
pd <- position_dodge(width=100)
westend.mean_dens_plot <- westend.meandens %>%
  filter(spp %in% c("mac","red","pter","ymac","purp","lam")) %>%
  mutate(type=ifelse(spp %in% c("mac","ymac","pter","lam"),"algae","urchin")) %>%
  ggplot(aes(x=date_str,y=meandens,col=long,linetype=type))+
      geom_hline(yintercept=0)+
      geom_line(lwd=2,alpha=0.8,position=pd)+
      geom_pointrange(aes(ymin=lower,ymax=upper),size=0.3,position=pd)+
      xlab("Year")+ylab("Normalized Density")+
      scale_color_manual(values=c("navyblue","gray50","darkgreen","darkcyan","mediumpurple4","darkred"),
                         guide=guide_legend(title="Species",override.aes = list(size=1)))+
      guides(linetype="none")+
      theme(text = element_text(color="black",size=16),
            legend.position = c(0.15,0.7),
            legend.text = element_text(color="black",size=10),
            panel.border = element_blank())
westend.mean_dens_plot
rm(pd,period.key,periods.first,westend.long,westend.meandens)

@@

@@

phys_effects_on_interactions[[1]]
phys_effects_on_interactions[[2]]
phys_effects_on_interactions[[3]]
phys_effects_on_interactions[[4]]
phys_effects_on_interactions[[5]]
phys_effects_on_interactions[[6]]

@@

### Effect of disturbance on *Macrocystis*
```{r disturbance and mac,fig.height=7,fig.width=5}
# with boxplot
quants <- c(0.05,0.5,0.95)
waves.mac.pter <- westend %>%
  select(one_of(c("pter","lam","mac","period"))) %>%
  distinct() %>%
  right_join(all.mod.coeffs.long,by="period")%>%
  gather("external_spp","ext_value",one_of(c('pter','lam','mac')))%>%
  filter(effect%in% c('waves','pter','lam'),mod=='mac')%>%
  ggplot(aes(ext_value,value))+
  geom_point(col='gray60',alpha=0.6)+
  geom_hline(yintercept = 0,linetype=2)+
  geom_quantile(quantiles=quants,lineend = "round",size=1,formula=y~x)+
  labs(x="Normalized Density",y="Effect of SWH on Macrocystis")+
  facet_grid(external_spp~effect,scales="free")
waves.mac.pter

```
## References

>>>>

theme_set(plot_theme)
```

@@

```{r mean density over time, echo=F,message=F}
# black plot theme (for presentations)
# for black plots
plot_theme_black <-   theme_minimal()+
  theme(text=element_text(family="Rockwell",size=12,color="white"),
        line = element_line(color="gray40"),
        legend.text = element_text(size=14),
        axis.title=element_text(family="Rockwell",size=14),
        axis.text=element_text(family="Rockwell",color="white",size=8),
        strip.background = element_rect(colour="black"),
        panel.border = element_rect(color="white",fill=NA),
        panel.background = element_rect(fill="black"),
        plot.background = element_rect(fill="black"),
        panel.grid = element_line(color="gray40"),
        panel.grid.major = element_line(color="gray40"),
        panel.grid.minor = element_line(color="gray40"))

# Species naming key for use in plotting
namekey <- tibble(dataset=c(rep("Benthic density",6),rep("Physical",5)),short=names(westend[3:13]),
                  long=c("Laminaria","Macrocystis >1m","Pterygophora","Purple urchin","Red urchin","Macrocystis <1m","Multivariate ENSO Index","Pacific Decadal Oscillation","North Pacific Gyre Oscillation","Significant Wave Height","Sea Surface Temperature"),plotting=c("L. far","M. pyr","P. cal","S. pur","M. fra","M. pyr (j)","(P) MEI","(P) PDO","(P) NPGO","(P) Max Hs","(P) SST"))

@@

# position adjustment for visual clarity
pd <- position_dodge(width=100)
westend.mean_dens_plot <- westend.meandens %>%
  filter(spp %in% c("mac","red","pter","ymac","purp","lam")) %>%
  mutate(type=ifelse(spp %in% c("mac","ymac","pter","lam"),"algae","urchin")) %>%
  ggplot(aes(x=date_str,y=meandens,col=long,linetype=type))+
      geom_hline(yintercept=0)+
      geom_line(lwd=2,alpha=0.8,position=pd)+
      geom_pointrange(aes(ymin=lower,ymax=upper),size=0.3,position=pd)+
      xlab("Year")+ylab("Normalized Density")+
      scale_color_manual(values=c("navyblue","gray50","darkgreen","darkcyan","mediumpurple4","darkred"),
                         guide=guide_legend(title="Species",override.aes = list(size=1)))+
      guides(linetype="none")+
      theme(text = element_text(color="black",size=16),
            legend.position = c(0.15,0.7),
            legend.text = element_text(color="black",size=10),
            panel.border = element_blank())
westend.mean_dens_plot
rm(pd,westend.long,westend.meandens)

```{r physical variables timeseries}
mei.pdo.npgo.ts.plot <-westend %>%
  ungroup()%>%
  slice(1:63)%>%
  select(period,mei,pdo,npgo)%>%
  gather("driver","value",mei:npgo,-period)%>%
  left_join(namekey, by=c("driver"="short"))%>%
  left_join(periods.first,by="period") %>%
  ggplot(aes(date_str,value,col=driver))+
  geom_line(lwd=1.5)+
  geom_hline(yintercept=0,linetype=2,col="black")+
  labs(title="Physical Drivers: Climate Modes",x="Year",y="Normalized Index Value")+
  plot_theme_black
sst.waves.ts.plot <-westend %>%
  ungroup()%>%
  slice(1:63)%>%
  select(period,sst,waves)%>%
  gather("driver","value",sst,waves,-period)%>%
  left_join(namekey, by=c("driver"="short"))%>%
  left_join(periods.first,by="period") %>%
  ggplot(aes(date_str,value,col=driver))+
  geom_line(lwd=1.5)+
  geom_hline(yintercept=0,linetype=2,col="black")+
  labs(title="Physical Drivers: Temperature and Disturbance",x="Year",y="Normalized Index Value")+
  plot_theme_black
mei.pdo.npgo.ts.plot
sst.waves.ts.plot
# ggsave("plots/climate_ts_plot.png",plot=mei.pdo.npgo.ts.plot,width=7,height=4)
# ggsave("plots/sst_waves_ts_plot.png",plot=sst.waves.ts.plot,width=7,height=4)

@@

### Example CCM

Each individual CCM analysis involves looks something like this:

```{r ccm example}
tempE <- westend.bestE['purp']
temp <- ccm(westend,lib=as.matrix(westend.segs),pred=as.matrix(westend.segs),E=tempE,lib_column= 'mac',target_column = 'purp',lib_sizes = c(10,25,50,75,100,125,150,200,300,400,500),num_samples=100,replace=T,silent=T,RNGseed = 41389)

mac_xmap_purp <- temp %>%
  group_by(lib_size)%>%
  summarise(rhomean=mean(rho,na.rm=T),upper=rhomean+sd(rho,na.rm=T),lower=rhomean-sd(rho,na.rm=T))%>%
  ungroup()%>%
  ggplot(aes(lib_size,rhomean))+
  geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.3,fill="red")+
  geom_line(color="darkorchid3")+
  labs(x="Library Size",y=expression(paste(rho, " (predictive skill)")),title="Macrocystis xmap Purple Urchin")

mac_xmap_purp
# ggsave("plots/ccm_example.png",plot=mac_xmap_purp,width=6,height=4)
```

@@

phys_effects_on_interactions[[1]]
phys_effects_on_interactions[[2]]
phys_effects_on_interactions[[3]]
phys_effects_on_interactions[[4]]
phys_effects_on_interactions[[5]]
phys_effects_on_interactions[[6]]
```

```{r some single plots}
# macrocystis, pterygophora, and temperature
quants<-c(0.05,0.5,0.95)
all.coeffs.with.phys %>%
  filter(type=="algal competition",mod%in% c('mac','pter'),effect %in% c('mac','pter'),external_forcing=="sst")%>%
  ggplot(aes(ext_value,value,col=factor(effect,levels=c("mac","ymac","pter","lam","purp","red"))))+
   geom_point(col='gray60',alpha=0.6)+
   geom_hline(yintercept = 0,linetype=2,col="white")+
   geom_quantile(quantiles=quants,lineend = "round",size=1,formula=y~x)+
   scale_color_manual(values=c('mac'="lightblue",'ymac'="gray30",'pter'="red",'lam'="darkcyan",'purp'="mediumpurple4",'red'="darkred"),guide="none")+
   labs(x="Normalized SST",y="Interaction Strength")+
   facet_grid(~mod,scales = "free_y")+
  plot_theme_black+
    theme(strip.text=element_blank(),
          plot.margin=unit(c(0.5, 0.5, 0.5, 0.5), units="line"))

# macrocystis, pterygophora, and NPGO
all.coeffs.with.phys %>%
  filter(type=="algal competition",mod%in% c('mac','pter'),effect %in% c('mac','pter'),external_forcing=="npgo")%>%
  ggplot(aes(ext_value,value,col=factor(effect,levels=c("mac","ymac","pter","lam","purp","red"))))+
   geom_point(col='gray60',alpha=0.6)+
   geom_hline(yintercept = 0,linetype=2,col="white")+
   geom_quantile(quantiles=quants,lineend = "round",size=1,formula=y~x)+
   scale_color_manual(values=c('mac'="lightblue",'ymac'="gray30",'pter'="red",'lam'="darkcyan",'purp'="mediumpurple4",'red'="darkred"),guide="none")+
   labs(x="Normalized NPGO",y="Interaction Strength")+
   facet_grid(~mod,scales = "free_y")+
  plot_theme_black+
    theme(strip.text=element_blank(),
          plot.margin=unit(c(0.5, 0.5, 0.5, 0.5), units="line"))

# macrocystis, pterygophora, herbivory, and NPGO
all.coeffs.with.phys %>%
  filter(type=="herbivory",mod%in% c('mac','pter'),effect %in% c('red','purp'),external_forcing=="npgo")%>%
  ggplot(aes(ext_value,value,col=factor(effect,levels=c("mac","ymac","pter","lam","purp","red"))))+
   geom_point(col='gray60',alpha=0.6)+
   geom_hline(yintercept = 0,linetype=2,col="white")+
   geom_quantile(quantiles=quants,lineend = "round",size=1,formula=y~x)+
   scale_color_manual(values=c('mac'="lightblue",'ymac'="gray30",'pter'="red",'lam'="darkcyan",'purp'="mediumpurple4",'red'="darkred"),guide="none")+
   labs(x="Normalized NPGO",y="Interaction Strength")+
   facet_grid(~mod,scales = "free_y")+
  plot_theme_black+
    theme(strip.text=element_blank(),
          plot.margin=unit(c(0.5, 0.5, 0.5, 0.5), units="line"))

# macrocystis, pterygophora, and waves
all.coeffs.with.phys %>%
  filter(type=="algal competition",mod%in% c('mac','pter'),effect %in% c('mac','pter'),external_forcing=="waves")%>%
  ggplot(aes(ext_value,value,col=factor(effect,levels=c("mac","ymac","pter","lam","purp","red"))))+
   geom_point(col='gray60',alpha=0.6)+
   geom_hline(yintercept = 0,linetype=2,col="white")+
   geom_quantile(quantiles=quants,lineend = "round",size=1,formula=y~x)+
   scale_color_manual(values=c('mac'="lightblue",'ymac'="gray30",'pter'="red",'lam'="darkcyan",'purp'="mediumpurple4",'red'="darkred"),guide="none")+
   labs(x="Normalized SWH",y="Interaction Strength")+
   facet_grid(~mod,scales = "free_y")+
  plot_theme_black+
    theme(strip.text=element_blank(),
          plot.margin=unit(c(0.5, 0.5, 0.5, 0.5), units="line"))

@@

### Effect of disturbance on *Macrocystis*
```{r disturbance and mac,fig.height=7,fig.width=5}
quants <- c(0.05,0.5,0.95)
waves.mac.pter <- westend %>%
  select(one_of(c("pter","lam","mac","period"))) %>%
  distinct() %>%
  right_join(all.mod.coeffs.long,by="period")%>%
  gather("external_spp","ext_value",one_of(c('pter','lam','mac')))%>%
  filter(effect%in% c('waves','pter','lam'),mod=='mac')%>%
  ggplot(aes(ext_value,value))+
  geom_point(col='gray60',alpha=0.6)+
  geom_hline(yintercept = 0,linetype=2)+
  geom_quantile(quantiles=quants,lineend = "round",size=1,formula=y~x)+
  labs(x="Normalized Density",y="Effect of SWH on Macrocystis")+
  facet_grid(external_spp~effect,scales="free")
waves.mac.pter
```

### Effect of Macrocystis on Laminaria/Pterygophora interaction
```{r disturbance and mac,fig.height=7,fig.width=5}
quants <- c(0.05,0.5,0.95)
mac.lam.pter <- westend %>%
  select(one_of(c("mac","period"))) %>%
  distinct() %>%
  right_join(all.mod.coeffs.long,by="period")%>%
  gather("external_spp","ext_value",mac)%>%
  filter(effect=='pter',mod=='lam')%>%
  ggplot(aes(ext_value,value))+
  geom_point(col='gray60',alpha=0.6)+
  geom_hline(yintercept = 0,linetype=2)+
  geom_quantile(quantiles=quants,lineend = "round",size=1,formula=y~x)+
  labs(x="Normalized Density",y="Effect of Pterygophora on Laminaria")
mac.lam.pter
```
## References
